---
title: "FunctionOperationVignette"
author: "Koshlan"
date: "August 18, 2015"
output: html_document
---

### A simple example of how to use *evaluate_polygon_contents()*


Suppose we have a simple raster object and a simple polygon object

We could use use raster::extract directly, but I have found that this can cause memory issue if too many pixels must be evaluated at once. So these wrapper offer an alternative
means to call raster::extract.

```{r}
require(raster)
# Load a raster and make it discrete values
X <- readRDS("Xraster.rds")
X@data@values <- floor(X@data@values*10) # Make it Discrete Integers
X2 <- X # Make a copy of the original raster
res(X2) <- c(5,5) # reduce the resolution, we lose any data values
X2p <- rasterToPolygons(X2) # make a 5x5 gride
V <-raster::extract(X, X2p, fun = modal) # Modal value inside each polygon
values(X2) <- unlist(V)
par(mfrow = c(1,2))
plot(X, main= "Original Raster")
plot(X2, main= "Extracted Modal Value")
plot(X2p, add= T)
head(unlist(V))
```

#### evaluate_polygon_contents()
For polygons here is a wrapper function specific to *nutrbutr*. It requires that the coordinate system is equivalent for both the polygon and raster file.

```{r}
require(sp)
devtools::load_all() # Loads nutrbutr
sp::proj4string(X) <- CRS("+proj=utm +zone=48 +datum=WGS84")
sp::proj4string(X2p) <- CRS("+proj=utm +zone=48 +datum=WGS84")
V <- nutrbutr::evaluate_polygon_contents(raster_input = X, polygon_input=X2p)
head(V)
```

#### evaluate_polygon_contents_by_chunks()
Here is another wrapper, which is useful for evaluating large files with many polygons, so as to avoid putting to so many pixel values into memory. 
```{r}
V <- evaluate_polygon_contents_by_chunks(raster_input = X,  polygon_input = X2p, chunk_size =25)
head(V)
```






